Constraints on the quasi-normal mode frequencies of the LIGO-Virgo signals by making full use of gravitational-wave modeling

### 2021-03-01 papers/intro_paper/results_lvc.tex
conclusions
more text about hierarchical method

Among all the GW signals detected so far, GW150914 (solid curve in Fig.~\ref{fig:o1o2_events}) is unique in its loudness, leading to the first, and to date, best attempt in measuring the QNM frequencies~\cite{TheLIGOScientific:2016src,Brito:2018rfr,Carullo:2019flw,Isi:2019aib}. Within 90\% credibility we obtain from GW150914: 
%
\begin{equation}
\df{220}=0.04^{+0.06}_{-0.04}\,\quad \dtau{220}=0.09^{+0.18}_{-0.18}\,.
\end{equation}

@@

As described in~\cite{Abbott:2020jks}, we can relax the assumption of a shared deviation across all events by using the hierarchical inference technique originally proposed in Refs.~\cite{Zimmerman:2019wzo,Isi:2019asy}. The general idea behind this technique is to assume that the non-GR parameters $(\df{220},\dtau{220})$ are drawn from a common underlying distribution, whose properties can be inferred from the population of events. Following~\cite{Zimmerman:2019wzo,Isi:2019asy,Abbott:2020jks} we model the population distribution with a Gaussian  $\mathcal{N}(\mu,\sigma)$ of unknown mean $\mu$ and standard deviation $\sigma$. We can then obtain a posterior distribution $P(\mu, \sigma |  \{d_j\}) $ for the hyperparameters $\mu$ and $\sigma$ from a joint analysis of all the GW events~\cite{Isi:2019asy}. We use the \texttt{stan}-based code~\cite{stan} developed and used in Refs.~\cite{Isi:2019asy,Abbott:2020jks} to obtain $P(\mu, \sigma |  \{d_j\}) $.
%
Finally, the distribution for a given non-GR parameter $\xi_{\text{nGR}}$ is then be obtained by computing~\cite{Isi:2019asy}
%
\begin{equation}
P(\xi_{\text{nGR}} | \{d_j\}) = \int P(\xi_{\text{nGR}} |\, \mu, \sigma)\,P(\mu, \sigma |  \{d_j\})\,d\mu\,d\sigma \,,
\end{equation}
%
where $P(\xi_{\text{nGR}} | \,\mu, \sigma) = \mathcal{N}(\mu,\sigma)$. The 1D posteriors for $\df{220}$ and $ \dtau{220}$ obtained through this technique are shown in Fig.~\ref{fig:o1o2_events} (dash-dotted curves) with corresponding median and 90\% credible interval given by: 
%
\begin{equation}
\df{220}=0.02^{+0.09}_{-0.09}\,,\quad \dtau{220}=0.13^{+0.42}_{-0.40}\,.
\end{equation}
Compared to Ref.~\cite{Abbott:2020jks} these constraints are almost a factor $\sim 4$ more stringent for $\df{220}$ and a factor $\sim 2$ for $\dtau{220}$. Similar improvements hold for the hyperparameters: $\df{220}\,(\mu=0.02^{+0.05}_{-0.05},\,\sigma<0.09)$ and $\dtau{220}\,(\mu=0.13^{+0.20}_{-0.18},\,\sigma<0.38)$.

>>>>

Among all the GW signals detected so far, GW150914 (solid curve in Fig.~\ref{fig:o1o2_events}) is unique in its loudness, leading to the first, and to date, best attempt in measuring the QNM frequencies~\cite{TheLIGOScientific:2016src,Brito:2018rfr,Carullo:2019flw,Isi:2019aib}. Within 90\% credibility we obtain from GW150914: 
%
\begin{equation}\label{GW150914_delta}
\df{220}=0.04^{+0.06}_{-0.04}\,\quad \dtau{220}=0.09^{+0.18}_{-0.18}\,.
\end{equation}

@@

As described in~\cite{Abbott:2020jks}, we can relax the assumption of a shared deviation across all events by using the hierarchical inference technique originally proposed in Refs.~\cite{Zimmerman:2019wzo,Isi:2019asy}. The general idea behind this technique is to assume that the non-GR parameters $(\df{220},\dtau{220})$ are drawn from a common underlying distribution, whose properties can be inferred from the population of events. Following~\cite{Zimmerman:2019wzo,Isi:2019asy,Abbott:2020jks} we model the population distribution with a Gaussian  $\mathcal{N}(\mu,\sigma)$ of unknown mean $\mu$ and standard deviation $\sigma$. Under those assumptions, the goal is then to directly measure a posterior distribution $P(\mu, \sigma |  \{d_j\}) $ for $\mu$ and $\sigma$ from a joint analysis of all the GW events. If GR is correct, then this posterior should be consistent with $\mu=0$ and $\sigma=0$. From Bayes' theorem it follows that~\cite{Isi:2019asy}:
%
\begin{equation}
P(\mu, \sigma |  \{d_j\}) \propto P(\mu,\sigma)\prod _{j=1}^N P(d_j |\, \mu, \sigma) \,,
\end{equation}
%
where $P(\mu,\sigma)$ is the prior (also known as hyperprior) on ($\mu,\sigma$), and $P(d_j | \mu, \sigma) $ can be written in terms of the individual likelihoods of a given non-GR parameter $\xi_{\text{nGR}}$ using~\cite{Isi:2019asy}
%
\begin{equation}
P(d_j | \mu, \sigma) = \int P(d_j |\, \xi_{\text{nGR}})  P(\xi_{\text{nGR}} |\, \mu, \sigma) \,d\xi_{\text{nGR}} \,.
\end{equation}
%
Here $P(\xi_{\text{nGR}} | \,\mu, \sigma) = \mathcal{N}(\mu,\sigma)$ by construction and $P(d_j |\, \xi_{\text{nGR}})$ is the likelihood for the parameter $\xi_{\text{nGR}}$ for a given event $d_j$ that is computed from the standard parameter estimation analysis. After obtaining $P(\mu, \sigma |  \{d_j\})$, we can then infer a population distribution for a given non-GR parameter $\xi_{\text{nGR}}$ using:~\cite{Isi:2019asy}
%
\begin{equation}\label{hier_combine}
P(\xi_{\text{nGR}} | \{d_j\}) = \int P(\xi_{\text{nGR}} |\, \mu, \sigma)\,P(\mu, \sigma |  \{d_j\})\,d\mu\,d\sigma \,.
\end{equation}
%
Notice that, if we fix $\sigma=0$, this approach is equivalent to assume that all events share the same non-GR parameter $\xi_{\text{nGR}} =\mu$ and~\eqref{hier_combine} reduces to the joint likelihood~\cite{Zimmerman:2019wzo}. In practice, we use the \texttt{stan}-based code~\cite{stan} developed and used in Refs.~\cite{Isi:2019asy,Abbott:2020jks} to obtain $P(\mu, \sigma |  \{d_j\}) $ and compute $P(\xi_{\text{nGR}} | \{d_j\})$.

The 1D posteriors for $\df{220}$ and $ \dtau{220}$ obtained through this technique are shown in Fig.~\ref{fig:o1o2_events} (dash-dotted curves) with corresponding median and 90\% credible interval given by: 
%
\begin{equation}
\df{220}=0.02^{+0.09}_{-0.09}\,,\quad \dtau{220}=0.13^{+0.42}_{-0.40}\,.
\end{equation}
Compared to Ref.~\cite{Abbott:2020jks} these constraints are almost a factor $\sim 4$ more stringent for $\df{220}$ and a factor $\sim 2$ for $\dtau{220}$. Similar improvements hold for the hyperparameters: $\df{220}\,(\mu=0.02^{+0.05}_{-0.05},\,\sigma<0.09)$ and $\dtau{220}\,(\mu=0.13^{+0.20}_{-0.18},\,\sigma<0.38)$.
